---
title: "BDA - Project"
author: "Anonymous"
output: 
  pdf_document: 
    toc: no
    toc_depth: 1
---

```{r setup, include=FALSE}
# This chunk just sets echo = TRUE as default (i.e. print all code)
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=TRUE, include=FALSE}
library(rstan)
library(ggplot2)
library(bayesplot)

options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)
Sys.setenv(LOCAL_CPPFLAGS = '-march=native')
SEED <- 48927 # set random seed for reproducability
```


# Analysis of survival chance relating to size of the maligant melanoma tumor size

## Introduction
This project analyses the chance to survive maligant melanoma with the thickness of melanoma tumor size. The data is a dataset collected at the Department of Plastic Surgery, University Hospital of Odense, Denmark from 1962 to 1977. Each patient had their tumour removed with surgical operation, with around 2,5cm of the surrounding skin. The dataset contains the survival time of each patient after the surgery, information about the patient (survived/died, sex, age), year of the operation, thickness of the tumour and if there were ulcers present in the tumour. The patients were followed to the end of 1977.

## Analysis

Read data.
```{r}
csv <- read.csv("../data/Melanoma.csv", header=TRUE)
data_melanoma <- as.data.frame(csv)
input <- list(N = nrow(data_melanoma),
              x = data_melanoma$time,
              y = data_melanoma$thickness)
```


### Quick glimpse inside the dataset

Now to give you a rough understanding of what the data set looks like:

```{r}
data_melanoma$ulcer <- as.logical(data_melanoma$ulcer)
data_melanoma$sex <- as.factor(data_melanoma$sex)

summary(data_melanoma)
head(data_melanoma)
```

### Scatterplot

Plot scatterplot with survival time and tumour thickness.
```{r}
ggplot() +
  geom_point(aes(x, y), data = data.frame(input), size = 1) +
  labs(y = 'Tumour thickness (cm)', x = 'Survival time (days)') +
  guides(linetype = F)
```


## Linear model

### Data

```{r}
writeLines(readLines("linear.stan"))
```

```{r}
fit_linear <- stan(file='linear.stan', data=input, seed=SEED)
print(fit_linear)
data_extract <- as.data.frame(fit_linear)
```

### Plot

Plotting to see if linear correlation can be found:

```{r}
plot(data_extract$alpha, data_extract$beta, xlab = "Alpha", ylab = "Beta")
abline(lm(data_extract$beta ~ data_extract$alpha), col = "red")
```

In the plot Alpha represents the time survived after the melanoma operation and Beta the thickness of the ulcer. The values seem to have a clear linear correlation in the way that those with smaller ulcers have lived longer after the operation.


### Chain converging

Check convergence of chains:
```{r}
rstan::traceplot(fit_linear, pars=c('alpha','beta'))
```

Although the scales are rather differing, the forms of the chains do behave a lot like one another.


### HMC diagnostics

HMC diagnostics:
```{r}
check_hmc_diagnostics(fit_linear)
```

None of the HMC-test output values give warnings, which can be considered a positive trait.


### Rhat

Rhat diagnostics:

```{r}
rhat(fit_linear)
```

The alpha and beta values are remarcably close to one another, meaning the chains are well converged.


### ESS -values

From monitor, the effective sample size ESS can be seen:

```{r}
monitor(fit_linear)
```


### Posterior densities
Plot posterior densities:

```{r}
stan_dens(fit_linear)
```



## Bernoulli model
Purpose of this model is to get information and predict the survival chance from melanoma, when the tumour thickness is known.

Clean data, 1 represents individuals that died because of melanoma, 0 represents individuals that are alive or died from other causes.
```{r}
data_melanoma[data_melanoma$status == 2,]$status <- 0
data_melanoma[data_melanoma$status == 3,]$status <- 0
```

Weakly informative priors were chosen for alpha and beta, both are normal(0, 10). The selected prior represents the thickness pretty well.
```{r}
writeLines(readLines("bernoulli.stan"))
```

Create input list and run the stan model.
```{r}
input_bernoulli <- list(N = nrow(data_melanoma),
                        x = data_melanoma$thickness,
                        y = data_melanoma$status)
fit_bernoulli <- stan(file='bernoulli.stan', data=input_bernoulli, seed=SEED, iter=5000, warmup=500)
print(fit_bernoulli)
```

### Chain converging

Check converge of chains:
```{r}
rstan::traceplot(fit_bernoulli, pars=c('alpha','beta'))
```

Chains seems to be pretty well converged.

### HMC diagnostics

HMC diagnostics:
```{r}
check_hmc_diagnostics(fit_bernoulli)
```

None of the values gives any warnings, so everything went well in the fitting.

### Rhat

Rhat diagnostics:
```{r}
rhat(fit_bernoulli)
```

Rhats are very close to one, so the chains are well converged.

### ESS -values

From monitor -function we can see the effective sample sizes (ESS -values).
```{r}
monitor(fit_bernoulli, warmup = 500)
```


### Posterior densities

Plot posterior densities:
```{r}
stan_dens(fit_bernoulli)
```

As can be seen from the beta, the expected probability of survival from melanoma is approximately 0.8 (chance of death is 0.2). It is dependant from the thickness of the tumour.


## Comparison of models

We used two different statistical models to analyze the data. Another was linear regression and another was bernoulli model. In these two models we approached the dataset in a bit different way. In linear regression we used variables: survival time and tumour thickness to predict the survival time when the tumour thickness is known. In bernoulli model we used status and thickness to predict the probability of death when the tumour thickness is known. They both measure the same thing in the end, which is survivability, but they address the problem in a different way and using partly different variables.


## Conclusion

These missing:
Posterior predictive checking
Predictive performance assessment if applicable (e.g. classification accuracy)
Sensitivity analysis with respect to prior choices


According to the data it can be concluded that the size of a tumor even if it is removed is in linear correlation with the patient having a shorter life span after the surgery. It can also be concluded that the size of the melanoma tumor has a correlation with the patient dying of cancer related reasons later on in life.

## Problems and potential improvements

The data set used was rather old having being made in the 60s and 70s. Also the data set was not that large making the results a bit less reliable. To improve this analysis we recommend taking more data points that correspond to the situation with current medical technology. Also the data should be gathered from a larger area than just one university hospital.

